---
title: "copy of gstore project"
author: "Jack Wright"
date: "11/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
```

##LOAD DATA

```{r}

train<-read.csv(here('kaggle_files','train_clean_jack.csv'))

target<-train%>%group_by(fullVisitorId)%>%summarize(total_revenue=sum(transactionRevenue))
```


```{r}
target$index<-as.numeric(row.names(target))
ggplot(target, aes(x=index,y=log(total_revenue)))+geom_point()
```


## visualization for users generating revenue (target=1)

```{r}
ggplot(target%>%mutate(rev_flag=case_when(total_revenue>0~TRUE,TRUE~FALSE)), aes(rev_flag))+geom_bar()
```

This shows that a very small percent of users generate the revenue

```{r}
temp_rev<-target%>%filter(total_revenue>0)

temp_rev%>%arrange(desc(total_revenue))%>%head()
```


## insights

-users generating rev use mac more than others
-channel groupiing is referral rather than organic
-use chrome
-visit more than once
-network domain is not unknown.known
-higher time on site
-total hits higher
-use desktop more
-mostly from US 
-usually connecting from CALI and NYC
-usually connect directly, less through google


## look at NaN values
```{r}
library(naniar)
gg_miss_var(train %>% filter(transactionRevenue > 0), show_pct = TRUE)
```

NOTE:

columns with more than 50% missing values. later remove them in data cleaning

it says newVisits only having nan and 1 as values is uninformative


## Constant columns

```{r}
uniq_counts<-train%>%
  select_if(negate(is.numeric))%>%
  apply(2,function(x) length(unique(x)))

names(uniq_counts[uniq_counts==1])
```
```{r}
unique(train$bounces)
```


## Defining the target

```{r}
y_train<-replace_na(train$transactionRevenue,0)
```


## Remove columms with missing values over a percentage

```{r}
train_select<-train%>%select_if(~mean(is.na(.))>.5)
```

```{r}
round(colMeans(is.na(train)),1)

```

