---
title: "Homework 1"
output: html_notebook
---


```{r}
library(tidyverse)
inc <-read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)
```

Preview of data

```{r}
summary(inc)
```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:


Summary of Industries for Fastest Growing industries
```{r}
inc<-inc%>%drop_na()
inc$Industry<-as.factor(inc$Industry)
nlevels(inc$Industry)
inc%>%group_by(Industry)%>%
  summarise(count_business=length(Industry))%>%arrange(desc(count_business))
```


Average number of employees per business in an industry
```{r}
inc%>%group_by(Industry)%>%
  summarise(avg_employee = as.integer(mean(Employees, na.rm=TRUE)))%>%arrange((desc(avg_employee)))
```


Total Employees per industry

```{r}
inc%>% group_by(Industry)%>%
  summarise(total_employee = mean(Employees)*length(Industry))%>%arrange(desc(total_employee))

```

Revenue per employee

```{r}
rev_emp<-inc%>% group_by(Industry)%>%
  summarise(total_employee = mean(Employees)*length(Industry), total_revenue=sum(Revenue), rev_per_emp = total_revenue/total_employee)%>%arrange(desc(rev_per_emp))
```

Looking at the head

```{r}
head(rev_emp)
```


I was expecting things like financial services and Insurance to be top performers in revenue per employee, but I wasn't expecting things like construction  and consumer products. I would think that employees would scale more directly with revenue here. 

Computer hardware and Construction stick out to me the most. I would expect things that needed to be built to require more employees to increase revenue. 

For computer hardware; since this is only looking at US businesses, maybe these are the companies that design and implement the products as opposed to actually manufacturing them, which I believe happens more overseas. 

Looking at the tail

```{r}
tail(rev_emp)
```


This makes a little more sense. Hospitality Education and Security are services provided by an employee directly without any additional product or service sold that generates revenue on its own. 

I am suprised to see engineering and Real Estate on this list, especially when construction is one of the highest revenue per employee categories. 



