---
title: "ch_8"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
```


## Logistic Regression


So far response variable is numeric and ideally follows a normal distribution


Now we look at where response variable is based on a series of yes or no responses, such as whether a particular restaurant is reccomeneded by being included in a prestegious guide. 


Ideally responses follow a binomial distribution in which case the appropriate model is a logeistic regression model


## Logistic Regression based on a single predictor


case of predicting a binomial random variable Y based on a single predictor variable x via logistic regression. 


### Binomial distribution


PROPERTIES:

1.

there are m identical trials

2.
each trial results in 2 ouctomes either success or failure

3.
$\theta$ the probability of 'success' is the same in all trials
4.
the trials are independent


these trials are bernoulli trials


Y= number of successes in m trials of a binomial process. Then Y is said to have a binomial distribution with parameters m and $\theta$

$$Y=Bin(m,\theta)$$


probability that Y takes the integer value of j (j=0,1,...m) is given by 

$$P(Y=j)= (m,j)\theta^j (1-\theta)^{m-j}= \frac{m!}{j!(m-j)!} \theta^j((m-j)!(1-\theta)^{m-j}$$


the mean and variance of Y are given by 

$$E(Y)=m\theta$$

$$Var(Y)=m\theta(1-\theta)$$



In logistic regression setting, we wish to model $\theta$ (probability of success) and hence Y, (because Y is the number of successes in m trials)


consider case of a single predictor variable x. 

i guess you look at the number of successes at each iteration (y_i) at each (m_i)

and 
1.

$y_i/m_i$ is an unbiased estimate of $\theta(x_i)$
2.

$y_i/m_i$ varies between 0 and 1


notice that the variance of response $y_i/m_i$ depends on $\theta(x_i)$ and theta is in the equation for variance, so it is not constant.  This means least squares regression is an innapropriate technique for analyzing Binomial responses. 


EX: Michelin and Zagat guides to NYC restaurants


Michelin is experts doing secret reviews

Zagats is mail in surveys



COMPARING TWO GUIDES for 164 french restaurants included in Zagat survey is ALSO included in Michelin guide



we want to model $\theta$, probability that a french restaurant from Zagats is included in Michelin based on the results of the survey


```{r}
file<-here('data','MichelinFood.txt')
df<-read.table(file,header = TRUE)
df
```

so `FOOD` is the food score, and also the x_i. So it looks like this dataframe has how many french restaurants at this food score are included. 

EX: 
```{r}
df[6,]
```

m_i the sample, has 18 restaurants. `y_i` the number of successses, (InMichelin when already in Zagats) 


`m_i-y_i`,  number of failures is NotInMichelin

`y_i/m_i` is the proportion , successes over trials. This is the estimate of $\theta(x_i)$ or the probability of success given the predictor variable x at a given level 


Look at theta

```{r}
ggplot(df, aes(x=Food,y=proportion))+geom_point()
```

the underlying function does not look linear, book says it looks `s` shaped...

low `Food` has almost 0 probablilty of success, but high `Food` has almost 1 probability



## Logistic function and odds

popular choice for the S shaped function is an exponential of the linear regression. it boils down to 

$$\beta_0+\beta_1 x = log(\frac{\theta(x)}{1-\theta(x)})$$


this quantity is called a `logit`

if this is a correct representation, plotting the logit vs the predictor should be a straight line

```{r}

df<-df%>%mutate(logit=log((proportion)/(1-proportion)))
ggplot(df, aes(x=proportion,y=logit))+geom_point()
```



the inner portion of the logit are the odds

$$\frac{\theta(x)}{1-\theta(x)}$$

concept of odds has 2 forms

*odds in favor* of success

-ratio of the probability of success over not success

*odds against* success
-inverse of this


bookies quote odds as odds against success. 

20 to 1 or 20 failures to 1 success in 21 races



BACK TO ZAGAT

x is the food rating for a restaurant

$\theta(x)$  is the probability of success (getting into Michelin)

logistic regression model for the response, (theta(x)) based on predictor x is given by

$$\theta(x)=\frac{1}{1+exp(-\{\beta_0+\beta_1 x\})}$$
```{r}
glm.model<-glm(formula=cbind(InMichelin,NotInMichelin)~Food,family=binomial, data=df)
summary(glm.model)
```

Plotting glm

```{r}
library(ggResidpanel)

resid_panel(glm.model,plots='all')

```
FOR PLOTTING:

need range of values of `Food` to produce fitted values



```{r}
range(df$Food)
#create sequence 
xFood<-15:28


```

use predict() function to create model for all the values of xweight

```{r}
yFood<-predict(glm.model, list(Food=xFood), type='response')
```

now plot
```{r}
ggplot(df, aes(x=Food, y=proportion))+geom_point()+geom_smooth(aes(x=xFood,y=yFood))
```
doing some math on the logit function you get

$$\frac{\hat{\theta}(x)}{1-\hat{\theta}(x)}=e^{\hat{\beta_0}+\hat{\beta_1}x}=e^{=10.842+.501x}$$

so thats what the odds OF SUCCESS will be equal to 
note that if the the food rating is increased by 

-1 unit, then the odds of being included in the Michelin increase by $e^{.501}$ or 1.7

-5 units, then odds of being included increase by $e^{5*5.01}$ or 12.2




## Likelihood for logistic Regression with a Single Predictor

plot of estimated probability and estimated odds from model

```{r}
x<-round(xFood,3)
theta<-round(yFood,3)
odds<-round(yFood/(1-yFood),3)

cbind(x,theta,odds)
```


Since we cant use Least squares( because variance is tied to theta in logistic regression ) we multiply together all of the probabilities of success to get *Likelihoood*


The Beta parameters can be estimated by maximizing log-likelihood. Has to be done with iteration


Standard approach to testing:

$$H_0:=\beta_1 =0$$


is to use the *Wald test Statistic*

$$Z=\frac{\hat{\beta_1}}{estimated~se(\hat{\beta_1})}$$


where the estimated standard error is calculated based on the iteratevely reweighted least squares approximation to the maximum likelihood estimate. 


The Wald test stat is then comared to a standard normal distribution to test for statistical significance. 


confidence intervals for the wald statistic:

$$\hat{\beta_1}\pm z_{1-\alpha/2}~estimated~se(\hat{\beta_1})$$





## Explanation of Deviance


residuals replaced by deviance in logistic regression

deviance is $G^2$

degrees of freedom in deviance

$$df=n-(number~of~\beta 's~estimated)$$


deviance in a logistic regression model (M) is comaring maximized log-likelihood under (M) with the maximized log likelihood under (S) (S is the saturated model that has a parameter for each observation)



### Saturated model

(S)

estimates theta by the observed proportion of 'successes' at $x_i$ 
-so i guess its if you made a theta for each observation explicitly?

heres what they say 

$$\hat{\theta_S} x(i)=y_i/m_i$$

So G^2

$$G^2=2[log(L_s)-log(L_M)]$$

when each m_i is large enough, the deviance can be used as a goodness-of-fit test for the logistic regression model


EX:

test if Zagats model is good enough

H_0: logistic regression model is appropriate

H_A: logistic model is innapropriate so a saturated model is needed

(kind of backwards? wierd)



```{r}
summary(glm.model)
```

looking at the residual deviance, the book says

$$P(G^2>11.368)=.498$$
```{r}

```

```{r}
pchisq(11.368,df=12, lower.tail = FALSE)
```
so we take the G^2 and look at it on a chisquare distribution to get the p-value


so we are unable to reject H_O.

The deviance goodness-of-fit test finds that the logistic regression model is an adequate fit overall for the Michelin Data



## Using Differences in Deviance Values to Compare Models


difference in deviance can be used to compare nested models. 

EX:

can compare null and residual deviances to test


differences between above example

$$G^2_{H_0}-G^2_{H_A}=61.427-11.268=50.059$$
This difference is to be compared to a $$\chi^2$$ distribution with $df_{H_O}-df_{H_A}=13-12=1$

```{r}
pchisq(50.059,df=1,lower.tail = FALSE)
```
OR

```{r}
anova(glm.model,update(glm.model,~1),test='Chisq')
```


earlier we found that the corresponding p-value based on the wall test (1.08e-8) Wald tests and tests based on the difference in deviances can result in quite different p-values



### R^2 for logistic regression


recall:


$$R^2=1-\frac{RSS}{SST}$$


RSS=sum of the squares of residuals of data from model

SST= total sum of squares, squares of difference from mean of data and the data points (like how different from 0 slope is it)


Since deviance in logistic regression  is a generalization of RSS 

$$R^2_dev = 1-\frac{G^2_{H_A}}{G^2_{H_0}}$$


for the single predictor Zagat model

$$R^2_dev = 1-\frac{11.368}{61.427}=.815$$


### Pearson goodness-of-fit statistic


like the deviance residuals, cant find out how to do them



## Binary Logistic Regression

Special case of logistic regression

-all the m =1 (only one trial for each)


goodness of fit for X^2 and G^2 are hard to interpret




EX:

Zagat survey again


consider each restaurant seperatly and classify each one according to wether it was included in michelin or not


y_i= 1 restaraunt InMichelin

y_i=0 restaurant not InMichelin

look at predictors:

-food, decor Service Price


```{r}
file<-here('data','MichelinNy.csv')
df<-read.csv(file, header=TRUE)
```

$\theta(x)$ probability restaurnt with Zagat food rating x is included in the Michelin guide. 


look at single predictor with binary response variable

```{r}
ggplot(df, aes(x=Food, y=InMichelin))+geom_jitter(height=.1,width=.1)
```
boxplot of food rating vs in michelin
```{r}
ggplot(df,aes(x=as.factor(InMichelin),y=Food))+geom_boxplot()
```


logistic regression

```{r}
summary(glm.model<-glm(formula=InMichelin~Food, family=binomial(),data=df))
```

```{r}
anova(glm.model,update(glm.model,~1),test='Chisq')
```

Note that the deviance and AIC values differ between the previous model and this binary model


## Deviance for the Case of Binary Data


for binary data, all the m are =1 so the saturated model S estimates \theta by the observed proportion of 'successes' at x_i

basically the deviance does not provide an assessment of the goodness-of-fit of the model when all m are =1

ALSO:

the deviance isn't a chisq distribution so that isn't useful either

ALTHOUGH:

the distribution of the DIFFERENCE of deviances is approximately a chisq distribution, so the Analysis of deviance table is valid



##  Residuals for Binary Data


## Transforming Predictors in Logistic Regression for Binary Data

circumstances under which the logistic regression model is appropriate for binary data and when to transform


The log odds equal to

$$log(\frac{\theta(x)}{1-\theta(x)})=log(\frac{P(Y=1)}{P(Y=0)}+log(\frac{f(x|Y=1)}{f(x|Y=0)}$$

where f(x|Y=j), j=0,1 is the *conditional density function* of the predictor given the value of the response

i think conditional density, is just a probability density function GIVEN the condition. (so they are normal distributions sum=1?)

since the first term doesn't change with X, we can ignore it

Suppose cdf = normal density with mean mu and variance sigma^2

MATH

when predictor variable x is normally distributed with different variances depending on if Y=1 or 0
**probably should copy latex from page 287 but im too tired**

some functions for values of \beta_0 1, and 2

when all of the variances between groups are the same

$$log(\frac{\theta(x)}{1-\theta(x)})=\beta_0+\beta_1 x$$


where 


$$\beta_1=\frac{\mu_1-\mu_0}{\sigma^2}$$
when predictor X is normally distributed with same variance for 2 values of Y, log odds are a linear function of x, with slope (beta_1)= to the difference in the mean of X across the two groups divided by common variance


